---
title: "p8131_stl2137_HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nnet)
```

## Part i
i) Summarize the data using appropriate tables of percentages to show the pair- wise associations between the levels of satisfaction and 1) contact with other residents and 2) type of housing. Comment on patterns in the associations.
```{r}
contact <- c(rep(c("tower block"), 2), rep(c("apartment"), 2), rep(c("house"), 2))

low_high_bin <- c(rep(c("low", "high"), 3))

low_satis <- c(65, 34, 130, 141, 67, 130)
medium_satis <- c(54, 47, 76, 116, 48, 105)
high_satis <- c(100, 100, 111, 191, 62, 104)

housing_dat <- cbind.data.frame(contact, low_high_bin, low_satis, medium_satis, high_satis)

housing_dat <- housing_dat %>% 
  mutate(
    total = dplyr::select(., low_satis:high_satis) %>% 
      rowSums(), 
    low_satis_pct = low_satis/total,
    mid_satis_pct = medium_satis/total,
    high_satis_pct = high_satis/total
  )
```

```{r, results='asis'}
knitr::kable(
  housing_dat, digits = 4
) 
```



## Part ii

ii) Use nominal logistic regression model for the associations between response variable, the levels of satisfaction, and the other two variables. 

Obtain a model that summarizes the patterns in the data. Describe your findings (the pattern in the associations, odds ratios with 95% confidence intervals, goodness of fit.) (Hint: use dummy variable for house types)

```{r}
### Nomial Logistic Regression
housing.nom <- multinom(cbind(low_satis, medium_satis, high_satis)~low_high_bin + contact)
summary(housing.nom)
```

```{r}
### Goodness of Fit
pihat = predict(housing.nom, type = 'probs')
m = rowSums(housing_dat[,3:5])
res.pearson = (housing_dat[,3:5]-pihat*m)/sqrt(pihat*m)
G.stat=sum(res.pearson^2) 
G.stat
# df = (n - p)(J - 1)
pval = 1 - pchisq(G.stat, df = (6-3)*(3-1))
```

Based off the p-value of `r pval`, we fail to reject the null hypothesis and conclude that the model fits the data. 

## Part iii

iii)  What does the ô°€tted model tell?

```{r}
housing_dat_long <- housing_dat %>% 
  mutate(
    contact = factor(contact, levels = c("apartment", "house", "tower block")),
    low_high_bin = factor(low_high_bin, levels = c("low", "high"))
  ) %>% 
  dplyr::select(contact, low_high_bin, low_satis, medium_satis, high_satis) %>% 
  pivot_longer(
    low_satis:high_satis,
    names_to = "satis_levels",
    values_to = "satis_num"
  ) %>% 
  mutate(
    satis_levels = factor(satis_levels, levels = c("low_satis", "medium_satis", "high_satis"), ordered = T)
  )
```

```{r}
library(MASS)
housing.polr <- polr(satis_levels ~ contact + low_high_bin, weights = satis_num, data = housing_dat_long)
summary(housing.polr)
```

The 

## Part iv

```{r}
pihat.ord = predict(housing.polr, housing_dat, type = 'p')
m.ord = rowSums(housing_dat[,3:5])
res.pearson.ord = (housing_dat[,3:5]-pihat.ord*m.ord)/sqrt(pihat.ord*m.ord)
G.ord = sum(res.pearson.ord^2)
G.ord 
```

The largest discrepancies between the observed frequencies and expected frequencies estimated from model are people with high satisfaction living in houses with high contact at a value of -1.478, people with medium satisfaction living in houses with high contact at a value of 1.368, and people with high satisfaction living in houses with low contact at a value 1.244. 