---
title: "p8131_stl2137_HW4"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nnet)
```

## Part i
```{r}
shelter <- c(rep(c("tower block"), 2), rep(c("apartment"), 2), rep(c("house"), 2))

contact <- c(rep(c("low", "high"), 3))

low_satis <- c(65, 34, 130, 141, 67, 130)
medium_satis <- c(54, 47, 76, 116, 48, 105)
high_satis <- c(100, 100, 111, 191, 62, 104)

housing_dat <- cbind.data.frame(shelter, contact, low_satis, medium_satis, high_satis) %>% 
    mutate(
    shelter = factor(shelter, levels = c("apartment", "house", "tower block")),
    contact = factor(contact, levels = c("low", "high"))
  ) 

housing_dat <- housing_dat %>% 
  mutate(
    total = dplyr::select(., low_satis:high_satis) %>% 
      rowSums(), 
    low_satis_pct = low_satis/total,
    mid_satis_pct = medium_satis/total,
    high_satis_pct = high_satis/total
  )
```

```{r, results='asis'}
knitr::kable(
  housing_dat, digits = 4
) 
```

* Those living in tower blocks for both low and high contact overall seem to have higher rates of high satisfaction compared to those living in apartments and houses.

* Those living in the tower block with high contact have the lowest percentage of low satisfaction and the highest percentage of high satisfaction. 

* For those living in apartments with low contact, the highest percentage of residents seem to have low satisfication; however, for those living in apartments with high contact, the highest percentage of residents seem to have high satisfication. 

* Those living in houses for both low and high contact have the highest percentage of low satisfication within people living in houses. 

## Part ii

```{r}
### Nomial Logistic Regression

housing.nom <- multinom(cbind(low_satis, medium_satis, high_satis)~shelter + contact, data = housing_dat)
summary_housing.nom <- summary(housing.nom)
summary_housing.nom
```

```{r}
### Confidence Intervals for Odds Ratios
### beta +/- + 1.96*se
exp_coef_housing <- exp(summary_housing.nom$coefficients)
coef_housing <- summary_housing.nom$coefficients
se_housing <- summary_housing.nom$standard.errors

ci_housing_plus <- exp(coef_housing + qnorm(0.025) * se_housing)
ci_housing_minus <- exp(coef_housing - qnorm(0.025) * se_housing)
```

### Low Satisfaction vs. Medium Satisfaction 

* The odds of having low satisfaction versus medium satisfaction living in an apartment with low contact is `r exp_coef_housing[1,1]`. We are 95% confident that the odds is between `r ci_housing_plus[1,1]` and `r ci_housing_minus[1,1]`.

* The odds ratio between living in an apartment and living in a house for low satisfaction versus medium satisfaction is `r exp_coef_housing[1,2]`, holding all other variables constant. We are 95% confident that the odds ratio is between `r ci_housing_plus[1,2]` and `r ci_housing_minus[1,2]`.

* The odds ratio between living in an apartment and living in a tower block for low satisfaction versus medium satisfaction is `r exp_coef_housing[1,3]`, holding all other variables constant. We are 95% confident that the odds ratio is between `r ci_housing_plus[1,3]` and `r ci_housing_minus[1,3]`.

* The odds ratio between living in housing with low contact versus high contact for low satisfaction versus medium satisfaction is `r exp_coef_housing[1,4]`, holding all other variables constant. We are 95% confident that the odds ratio is between `r ci_housing_plus[1,4]` and `r ci_housing_minus[1,4]`.

### Low Satisfaction vs. High Satisfaction 

* The odds of having low satisfaction versus high satisfaction living in an apartment with low contact is `r exp_coef_housing[2,1]`. We are 95% confident that the odds is between `r ci_housing_plus[2,1]` and `r ci_housing_minus[2,1]`.

* The odds ratio between living in an apartment and living in a house for low satisfaction versus high satisfaction is `r exp_coef_housing[2,2]`, holding all other variables constant. We are 95% confident that the odds ratio is between `r ci_housing_plus[2,2]` and `r ci_housing_minus[2,2]`.

* The odds ratio between living in an apartment and living in a tower block for low satisfaction versus high satisfaction is `r exp_coef_housing[2,3]`, holding all other variables constant. We are 95% confident that the odds ratio is between `r ci_housing_plus[2,3]` and `r ci_housing_minus[2,3]`.

* The odds ratio between living in housing with low contact versus high contact for low satisfaction versus high satisfaction is `r exp_coef_housing[2,4]`, holding all other variables constant. We are 95% confident that the odds ratio is between `r ci_housing_plus[2,4]` and `r ci_housing_minus[2,4]`.

```{r}
### Goodness of Fit
pihat = predict(housing.nom, type = 'probs')
m = rowSums(housing_dat[,3:5])
res.pearson = (housing_dat[,3:5]-pihat*m)/sqrt(pihat*m)
G.stat=sum(res.pearson^2) 
G.stat
# df = (n - p)(J - 1)
pval = 1 - pchisq(G.stat, df = (6-4)*(3-1))
```

Based off the p-value of `r pval`, we fail to reject the null hypothesis and conclude that the model fits the data. 

## Part iii

```{r}
housing_dat_long <- housing_dat %>% 
  dplyr::select(shelter, contact, low_satis, medium_satis, high_satis) %>% 
  pivot_longer(
    low_satis:high_satis,
    names_to = "satis_levels",
    values_to = "satis_num"
  ) %>% 
  mutate(
    satis_levels = factor(satis_levels, levels = c("low_satis", "medium_satis", "high_satis"), ordered = T)
  )
```

```{r}
library(MASS)
housing.polr <- polr(satis_levels ~ shelter + contact, weights = satis_num, data = housing_dat_long)
summary(housing.polr)
```

* 0.2353 is the log odds ratio of falling into the lower category of satisfaction versus the higher category of satisfaction for living in an apartment versus living in a house, holding all other variables constant.

* -0.5010 is the log odds ratio of falling into the lower category of satisfaction versus the higher category of satisfaction for living in an apartment versus living in a tower block, holding all other variables constant. 

* -0.2524 is the log odds ratio of falling into the lower category of satisfaction versus the higher category of satisfaction for living in housing with low contact versus living in housing with high contact, holding all other variables constant. 

## Part iv

```{r}
pihat.ord = predict(housing.polr, housing_dat, type = 'p')
m.ord = rowSums(housing_dat[,3:5])
res.pearson.ord = (housing_dat[,3:5]-pihat.ord*m.ord)/sqrt(pihat.ord*m.ord)
G.ord = sum(res.pearson.ord^2)
G.ord 
```

The largest discrepancies between the observed frequencies and expected frequencies estimated from model are people with high satisfaction living in houses with high contact at a value of -1.478, people with medium satisfaction living in houses with high contact at a value of 1.368, and people with high satisfaction living in houses with low contact at a value 1.244. 